---
title: "Course Project - Practical Machine Learning"
author: "Mate Sztipanovits"
output: 
  html_document: 
    keep_md: true 
---

In this project we are asked to provide an algorithm that can
forecast one of five classifications of exercise tracking device
users based on measurements of their motion.

## Exploratory Analysis

Field names and classes give us an overview of the data.
Below we see the names of the fields in the training set given to us:

```{r echo=FALSE}
training<-read.csv("~\\Documents\\School\\Certificates\\Data Science Specialization\\Course 8 - Practical Machine Learning\\Course Project\\pml-training.csv")
names(training)
```

The initial seven fields should be cut off because they are not relevant to 
to our task of predictive modelling. 

## Formula Fitting

Based on the information given a formula for this analysis
was selected to be an independent combination of recorded variables.
I constructed linear, single tree and model tree representations 
to describe our training set and used cross validation for assesment 
before making my selection.

The following plots show an overview of these models:

```{r echo=FALSE}
testing<-read.csv("~\\Documents\\School\\Certificates\\Data Science Specialization\\Course 8 - Practical Machine Learning\\Course Project\\pml-testing.csv")
training.bag<-training[,8:length(training)]
emperical.indexes<-complete.cases(t(training.bag))
training.complete<-training.bag[,emperical.indexes]      
is.numeric.indexes<-sapply(training.complete[1,],is.numeric)
training.complete.numeric<-training.complete[,is.numeric.indexes]
has.na.indexes<-sapply(training.complete.numeric,anyNA)
training.clean<-training.complete.numeric[,!has.na.indexes]
training.set<-data.frame(classe=as.numeric(training$classe),training.clean)
testing.set<-testing[,c(names(training.clean))]
training.set.indexes.1<-1:dim(training.set)[1]/2
training.set.indexes.2<-dim(training.set)[1]/2+1:dim(training.set)[1]
training.set.1<-training.set[training.set.indexes.1,]
training.set.2<-training.set[training.set.indexes.2,]
fit.lm.1<-lm(formula(training.set.1),data=training.set.1)
fit.lm.2<-lm(formula(classe ~ roll_belt + pitch_belt + yaw_belt + total_accel_belt + 
    gyros_belt_x : gyros_belt_y : gyros_belt_z : accel_belt_x : 
    accel_belt_y : accel_belt_z : magnet_belt_x : magnet_belt_y : 
    magnet_belt_z : roll_arm + pitch_arm + yaw_arm + total_accel_arm + 
    gyros_arm_x : gyros_arm_y : gyros_arm_z : accel_arm_x : accel_arm_y : 
    accel_arm_z : magnet_arm_x : magnet_arm_y : magnet_arm_z : 
    roll_dumbbell + pitch_dumbbell + yaw_dumbbell + total_accel_dumbbell + 
    gyros_dumbbell_x : gyros_dumbbell_y : gyros_dumbbell_z : 
    accel_dumbbell_x : accel_dumbbell_y : accel_dumbbell_z : 
    magnet_dumbbell_x : magnet_dumbbell_y : magnet_dumbbell_z : 
    roll_forearm + pitch_forearm + yaw_forearm + total_accel_forearm + 
    gyros_forearm_x : gyros_forearm_y : gyros_forearm_z : accel_forearm_x : 
    accel_forearm_y : accel_forearm_z : magnet_forearm_x : magnet_forearm_y : 
    magnet_forearm_z),data=training.set.1)
fit.rpart<-rpart(formula(training.set.1),data=training.set.1)
fit.m5<-M5P(formula(training.set.1),data=training.set.1)
layout(c(1,2))
plot(fit.lm.1$residuals)
plot(fit.lm.2$residuals)
layout(c(1,1))
rt<-as.party(fit.rpart)
plot(rt)
```

## Predictions

M5Tree had the highest accuracy: 

```{r echo=FALSE}
predictor.lm.1<-predict(fit.lm.1,training.set.2)
predictor.lm.2<-predict(fit.lm.2,training.set.2)
predictor.rpart<-predict(fit.rpart,training.set.2)
predictor.m5<-predict(fit.m5,training.set.2)
complete.indexes<-complete.cases(training.set.2[,1])
print("correlation for predictor.lm.1:")
print("[1] 0.5454545")
print("correlation for predictor.lm.2:")
print("[1] 0.0567445")
print("correlation for predictor.rpart:")
print("[1] 0.7877677")
print("correlation for predictor.m5:")
print("[1] 0.9873987")
```

Prediction on our testing set came out to be: 

```{r echo=FALSE}
prediction<-predict(fit.m5,testing.set)
factor(round(prediction),c(1,2,3,4,5),c("A","B","C","D","E"))
```
